---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
editor: 
  markdown: 
    wrap: 72
---

### Installing dslabs package

install.packages("dslabs") ###load packages library("dslabs")
library("dplyr") library("ggplot2") ###look at help file for gapminder
data help(gapminder) ###get an overview of data structure str(gapminder)
###get a summary of data summary(gapminder) #determine the type of
object gapminder is class(gapminder)

### Let's view the curent data

View(gapminder)

###Creating a dataset for only African countries africadata \<-
gapminder %\>% filter(continent == "Africa")

### View the africadata created

View(africadata)

###Get an overview of data strucure str(africadata)

###Get a summary of data summary(africadata)

### Create a new object from africa data with only infant_mortality and life_expectancy

africadata_v1 \<- africadata\[,3:4\]

###Get an overview of data structure str(africadata_v1)

### Get summary of data

summary(africadata_v1)

### Create another object from africadata with only population and life expectancy

africadata_v2 \<- africadata\[,c(6,4)\]

### Get an overview of data structure

str(africadata_v2)

### Summary of data

summary(africadata_v2)

### Plot infant mortality and life expectancy

ggplot(africadata_v1, aes(x=life_expectancy, y = infant_mortality)) +
geom_point()

### Plot infant mortality and Population size

ggplot(africadata_v2, aes(x=life_expectancy, y = population)) +
geom_point()

### Getting the missing records for infant mortality

is.na(africadata\$infant_mortality)

### New object excluding missing values

africadata_2000 \<- africadata %\>% filter(year == 2000)

### Check structure of the data

str(africadata_2000)

### Get a summary of the data

summary(africadata_2000)

### Ploting life expectancy and infant mortality for the year 2000

ggplot(africadata_2000, aes(x=life_expectancy, y = infant_mortality)) +
geom_point()

### Ploting life expectancy and population size for the year 2000

ggplot(africadata_2000, aes(x=life_expectancy, y = population)) +
geom_point()

### Fitting a linear regression: using infant mortality as a predictor

fit1 \<- lm(life_expectancy \~ infant_mortality, data = africadata_2000)
\### Print result summary(fit1)

### Using population size as a predictor

fit2 \<- lm(life_expectancy \~ population, data = africadata_2000) \###
Print result summary(fit2)

# Kelly Hatfield's Contributions

This section is added by Kelly Hatfield.

## Assessing the Impact of region in the relationship between life expectancy and mortality in year 2000

First, we will recreate the visualization of the Africa data in 2000 but
adding in color to differentiate the differences in region.

```{r}
#load packages

library(dslabs) 
library("dplyr") 
library("ggplot2") 

africadata <- gapminder %>% filter(continent == "Africa")
#reset africa dataset for year 2000 only
africadata_2000 = subset(africadata, year==2000)

#More Plotting #Infant Mortality and Life Expectancy in 2000;
ggplot(africadata_2000, aes(x=infant_mortality, y=life_expectancy, color=region)) +geom_point()


```

Next we will run a simple linear regression model to assess the mean
life expectancy in each of the regions

```{r}

#Box Plots of Life Expectancy by Region for the years 2000;

box=ggplot(africadata_2000, aes(x=region, y=life_expectancy, color=region)) + geom_boxplot()

box + geom_jitter(shape=16, position=position_jitter(0.2))

#Simple linear regression model for year 2000; 

fit3 = lm (life_expectancy~factor(region), data=africadata_2000)

summary(fit3)

```

This shows us that there are significant variability in the mean life
expectancy by region in the year 2000.

Next, we will assess region and infant mortality in the same model.

```{r}

#MLR
adjfit =lm(life_expectancy~infant_mortality+factor(region), data=africadata_2000)

summary(adjfit)

```

From this model we can interpret that when adjusting for region in
Africa, infant mortality is still linearly associated with life
expectancy (p\<0.001). The parameter estimate tells us that for each
one-unit increase in infant mortality life expectancy declines by 0.18
years.
