---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
editor: 
  markdown: 
    wrap: 72
---


### Loading libraries
```{r}
library("dslabs")
library("dplyr") 
library("ggplot2")
```

###look at help file for gapminder data
```{r}
help(gapminder)
```
 
###get an overview of data structure 
```{r}
str(gapminder)
```


###get a summary of data
```{r}
summary(gapminder)
```


### determine the type of object gapminder is 
```{r}
class(gapminder)
```


###Creating a dataset for only African countries
```{r}
africadata <- gapminder %>% filter(continent == "Africa")
```


###Get an overview of data strucure
```{r}
str(africadata)
```

###Get a summary of data 
```{r}
summary(africadata)
```



### Create a new object from africa data with only infant_mortality and life_expectancy
```{r}
africadata_v1 <- africadata[,3:4]
```


###Get an overview of data structure
```{r}
str(africadata_v1)
```


### Get summary of data
```{r}
summary(africadata_v1)
```

View(africadata)
### Create another object from africadata with only population and life expectancy
```{r}
africadata_v2 <- africadata[,c(4,6)]
```

### Get an overview of data structure
```{r}
str(africadata_v2)
```

### Summary of data
```{r}
summary(africadata_v2)
```



### Plot infant mortality and life expectancy
```{r}
ggplot(africadata_v1, aes(x=life_expectancy, y = infant_mortality)) +
geom_point()
```


### Plot infant mortality and Population size
```{r}
ggplot(africadata_v2, aes(x=life_expectancy, y = population)) +
geom_point()

```


### Getting the missing records for infant mortality
```{r}
is.na(africadata$infant_mortality)
```


### New object excluding missing values
```{r}
africadata_2000 <- africadata %>% filter(year == 2000)
```


### Check structure of the data
```{r}
str(africadata_2000)
```


### Get a summary of the data
```{r}
summary(africadata_2000)
```


### Ploting life expectancy and infant mortality for the year 2000
```{r}
ggplot(africadata_2000, aes(x=life_expectancy, y = infant_mortality)) +
geom_point()
```


### Ploting life expectancy and population size for the year 2000
```{r}
ggplot(africadata_2000, aes(x=life_expectancy, y = population)) +
geom_point()
```


### Fitting a linear regression: using infant mortality as a predictor
```{r}
fit1 <- lm(life_expectancy ~ infant_mortality, data = africadata_2000)
```

### Print result 
```{r}
summary(fit1)
```


### Using population size as a predictor
```{r}
fit2 <- lm(life_expectancy ~ population, data = africadata_2000)
```


###Print result 
```{r}
summary(fit2)
```



# Kelly Hatfield's Contributions

This section is added by Kelly Hatfield.

## Assessing the Impact of region in the relationship between life expectancy and mortality in year 2000

First, we will recreate the visualization of the Africa data in 2000 but
adding in color to differentiate the differences in region.

```{r}
#load packages

library(dslabs) 
library("dplyr") 
library("ggplot2") 

africadata <- gapminder %>% filter(continent == "Africa")
#reset africa dataset for year 2000 only
africadata_2000 = subset(africadata, year==2000)

#More Plotting #Infant Mortality and Life Expectancy in 2000;
ggplot(africadata_2000, aes(x=infant_mortality, y=life_expectancy, color=region)) +geom_point()


```

Next we will run a simple linear regression model to assess the mean
life expectancy in each of the regions

```{r}

#Box Plots of Life Expectancy by Region for the years 2000;

box=ggplot(africadata_2000, aes(x=region, y=life_expectancy, color=region)) + geom_boxplot()

box + geom_jitter(shape=16, position=position_jitter(0.2))

#Simple linear regression model for year 2000; 

fit3 = lm (life_expectancy~factor(region), data=africadata_2000)

summary(fit3)

```

This shows us that there are significant variability in the mean life
expectancy by region in the year 2000.

Next, we will assess region and infant mortality in the same model.

```{r}

#MLR
adjfit =lm(life_expectancy~infant_mortality+factor(region), data=africadata_2000)

summary(adjfit)

```

From this model we can interpret that when adjusting for region in
Africa, infant mortality is still linearly associated with life
expectancy (p\<0.001). The parameter estimate tells us that for each
one-unit increase in infant mortality life expectancy declines by 0.18
years.
