---
title: "R Data Wrangling Exercise - Chris Okitondo"
output:
  html_document:
    toc: FALSE
---



This dataset shows health conditions and contributing causes mentioned in conjunction with deaths involving coronavirus disease 2019 (COVID-19) by age group and jurisdiction of occurrence from 2020-2023. It was downloaded from the CDC website.

# Setup
Load needed packages. make sure they are installed.

```{r, echo=FALSE, message=FALSE}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(ggplot2) # For data visualization
library("tinytex")
```


# Data loading
```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location<- here::here("data","raw_data","Conditions_Contributing_to_COVID-19_Deaths_2020_2023.xlsx")
rawdata <- readxl::read_excel(data_location)
```




# Looking at the data
```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```
# Cleaning

# Telling R to avoid using scientific notation
```{r}
options(scipen=999)
```


# List variable names
```{r}
names(rawdata)
```

# Let's rename some variable names
```{r}
rawdata <- rawdata %>% rename(current_data = `Data As Of`, start_date = `Start Date`, end_date = `End Date`, condition_group = `Condition Group`, condition = Condition, age_group = `Age Group`, covid_deaths = `COVID-19 Deaths`)
```


# Distribution of age group
```{r}
table(rawdata$age_group)
```

# Excluding data where age group is not stated
```{r}
rawdata <- rawdata %>% filter(age_group != 'Not stated')
```
Observation went from 521640 to 469476



# Let's visualize at COVID deaths by age group
```{r}
ggplot(data = rawdata, aes(x = age_group, y = covid_deaths)) + geom_bar(stat='identity', fill = "steelblue")
```

# Saving the cleaned data
```{r}
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(rawdata, file = save_data_location)
```





##This Section Added by Annabella Hines

```{r}
cleandata <- readRDS(file=save_data_location)
```



##Plots

First I wanted to look at the number of Covid deaths by each condition group.
```{r}
cleandata2 <- cleandata %>% filter(Group=="By Total")
ggplot(cleandata2, aes(x=condition_group, y=covid_deaths))+geom_col()+xlab("Condition Group")+ylab("Covid Deaths")+ theme(axis.text.x = element_text(angle=13, size=5))
```
It appears that respiratory diseases had the highest number of associated Covid deaths. I now want to look at the respiratory disease by specific condition.
```{r}
respiratory <- cleandata2 %>% filter(condition_group=="Respiratory diseases")
ggplot(respiratory, aes(x=condition, y=covid_deaths))+geom_col()+xlab("Respiratory Condition")+ylab("Covid Deaths")+theme(axis.text.x = element_text(angle=13))
```
The respiratory condition with the highest number of Covid deaths appears to be the influenza and pneumonia.
The last thing I want to do is look as the distribution of respiratory diseases by age group.
```{r}
ggplot(respiratory, aes(x=age_group, y=covid_deaths, fill=condition))+geom_col()+theme(legend.key.size = unit(0.25, 'cm'), axis.text.x = element_text(angle=90))+ylab("Covid Deaths")+ xlab("Age Group")
```


