---
title: "Data Exploratory Analysis"
author: "Chris Okitondo"
date: "03/03/2023"
output: html_document
---

### Setup
```{r, warning=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyverse)
library(reshape2)
```


### Loading data
```{r}
#Path to data. 
data_location <- here::here("fluanalysis", "data", "cleaned_data.rds")
#load data
mydata <- readRDS(data_location)
```


### Looking at the data
```{r}
dplyr::glimpse(mydata)
summary(mydata)
head(mydata)
skimr::skim(mydata)
```

### Checking variable names in the dataset
```{r}
names(mydata)
```

## Continuous outcome variable: Body Temperature
### Looking at the distribution of body temperature, our continuous outcome variable
```{r, warning=FALSE, message=FALSE}
ggplot(mydata, aes(x = BodyTemp)) +
  geom_histogram() + ggtitle("Distribution of Body Temperature") + theme_classic()
```
### Creating boxplot of Cough intensity and Body temperature variables
```{r}
# Creating boxplot of Cough intensity and Body temperature variables
ggplot(mydata, aes(x = factor(CoughIntensity), y = BodyTemp)) +
  geom_boxplot() +
  labs(x = "Cough Intensity", y = "Body Temperature") + ggtitle("Patterns between cough intensity abd body temperature") + theme_classic()
```

### Creating boxplot of Chest Congestion and Body temperature variables
```{r, message=FALSE, warning=FALSE}
ggplot(mydata, aes(x = factor(ChestCongestion), y = BodyTemp)) +
  geom_boxplot() +
  labs(x = "Chest Congestion", y = "Body Temperature") + ggtitle("Patterns between ChestCongestion and body temperature") + theme_classic()
```

### Creating boxplot of Headache and Body temperature variables
```{r, message=FALSE, warning=FALSE}
# Creating boxplot of Cough intensity and Body temperature variables
ggplot(mydata, aes(x = factor(Headache), y = BodyTemp)) +
  geom_boxplot() +
  labs(x = "Headache", y = "Body Temperature") + ggtitle("Patterns between Headache and body temperature") + theme_classic()
```

## Categorical outcome variable
### The distribution of Nausea
```{r, warning=FALSE, message=FALSE}
ggplot(mydata, aes(x = Nausea)) +
  geom_bar() + ggtitle("Distribution of Nausea") + theme_classic()
```

### Distribution of Nausea and Nasal Congestion
```{r, message=FALSE, warning=FALSE}
ggplot(mydata, aes(x = Nausea, fill = NasalCongestion)) +
  geom_histogram(stat="count") + 
  ggtitle("Distribution of Nausea and Nasal Congestion") + theme_classic()
```

### Distribution of Nausea and Nasal Congestion
```{r, message=FALSE, warning=FALSE}
ggplot(mydata, aes(x = Nausea, fill = Fatigue)) +
  geom_histogram(stat="count") + 
  ggtitle("Distribution of Nausea and Fatigue") + theme_classic()
```

