---
title: "Model Fitting"
author: "Chris Okitondo"
date: "03/03/2023"
output: html_document
---


In this section, we will explore the following tasks:
1. Loads cleaned data.
2. Fits a linear model to the continuous outcome (Body temperature) using only the main predictor of interest.
3. Fits another linear model to the continuous outcome using all (important) predictors of interest.
4. Compares the model results for the model with just the main predictor and all predictors.
5. Fits a logistic model to the categorical outcome (Nausea) using only the main predictor of interest.
6. Fits another logistic model to the categorical outcome using all (important) predictors of interest.
7. Compares the model results for the categorical model with just the main predictor and all predictors.

### Setup
```{r, warning=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(performance)
```



### Loading data
```{r}
#Path to data. 
data_location <- here::here("fluanalysis", "data", "cleaned_data.rds")
#load data
mydata <- readRDS(data_location)
```


### This sets up the linear model
```{r}
lm_mod <- linear_reg() %>% set_engine("lm")
```


### Fitting a linear model to the continuous outcome (Body temp) using only the main predictor of interest (RunnyNose)
```{r}
lm_fit1 <- lm_mod %>% fit(BodyTemp ~ RunnyNose, data = mydata)
tidy(lm_fit1)
```
### Fitting a linear model to the continuous outcome (Body temp) using all other predictors of interest
```{r}
lm_fit2 <- lm_mod %>% fit(BodyTemp ~ ., data = mydata)
tidy(lm_fit2)
```
### Comparing the above linear models: lm_fit1 which is a reduced model and lm_fit2 which is a full model
```{r}
anova(lm_fit1$fit, lm_fit2$fit)
glance(lm_fit1) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)
glance(lm_fit2) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)
# Another way to compare these models
compare_performance(lm_fit1, lm_fit2)
```

### This sets up logistics regression
```{r}
log_mod <- logistic_reg() %>%
  set_engine("glm")
```


### Fitting logistic model to the categorical outcome (Nausea) using only the main predictor of interest (RunnyNose)
```{r}
log_fit1 <- log_mod %>% fit(Nausea ~ RunnyNose, data = mydata)
tidy(log_fit1)
```
### Fitting logistic model to the categorical outcome (Nausea) using all other predictors of interest 
```{r}
log_fit2 <- log_mod %>% fit(Nausea ~ ., data = mydata)
tidy(log_fit2)
```

### Comparing the above logistic models: log_fit1 which is a reduced model and log_fit2 which is a full model
```{r}
anova(log_fit1$fit, log_fit2$fit)
compare_performance(log_fit1, log_fit2)
```


