---
title: "Model Evaluation"
author: "Chris Okitondo"
date: "03/22/2023"
output: html_document
---

### Loading necessary packages
```{r, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(tidymodels)
library(performance)
library(dplyr)
library(skimr)
```


### Loading previously processed data
```{r, message=FALSE, warning=FALSE}
#Path to data. 
data_location <- here::here("fluanalysis", "data", "cleaned_data.rds")
#load data
mydata <- readRDS(data_location)
```


### Data splitting
* Here, we are splitting our data into two subsets. 3/4 for the training and 1/4 for test.
```{r}
# To maintain reproducible results when re-done.
set.seed(1234) 

# Proceeding with 75% for training and 25% for test data
split <- initial_split(mydata, prop = 3/4)
```

#### Dataframe for the training and test data
```{r}
train_data <- training(split)
test_data  <- testing(split)
```

* Training data set now contains 547 observations while test data contains 183.

### Workflow creation and model fitting

#### FULL MODEL: Creating recipe that fits a logistic model with nausea as the outcome and all the predictors in the data 

```{r, warning=FALSE, message=FALSE}
# Outcome is nausea. The rest of variables are all predictors
# Recipe #1: Nausea predicted by all variables
flu_mod10_rec <- recipe(Nausea ~ ., data = train_data)

# Model: logistic regression using GLM engine
flu_mod10_mod <- logistic_reg() %>% 
                    set_engine("glm")

# Workflow: Pairing model and recipe 
flu_mod10_workflow <- workflow() %>% 
  add_model(flu_mod10_mod) %>% 
  add_recipe(flu_mod10_rec)
# Pring workflow
flu_mod10_workflow

# Fitting the model to a training dataset
flu_mod10_fit <- 
 flu_mod10_workflow %>% 
  fit(data = train_data)

# Looking at the model output
flu_mod10_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

#### FULL MODEL: USING trained workflow to predict unseen test data

```{r, warning = FALSE, message=FALSE}
# Using the trained workflow (flu_mod10_fit) to predict with the unseen test data
predict(flu_mod10_fit, test_data)

# Using argument() with the model plus test data for saving them together
flu_mod10_aug <- 
  augment(flu_mod10_fit, test_data)
  
flu_mod10_aug %>%
  select(Nausea, .pred_No, .pred_Yes)
  
# Plotting ROC curve
flu_mod10_aug %>% 
  roc_curve(truth = Nausea, .pred_No) %>% 
  autoplot()

# Using roc_auc() to get the estimates
flu_mod10_aug %>% 
  roc_auc(truth = Nausea, .pred_No)
```


#### ALTERNATIVE MODEL: Outcome is still nausea and predictor is runnynose 
```{r, warning=FALSE, message=FALSE}
# Using Nausea as a categorical outcome of interest and RunnyNose as main predictor
flu_mod10_rec2 <- recipe(Nausea ~ RunnyNose, data = train_data)

# Fitting the logistic model
flu_mod10_mod2 <- logistic_reg() %>% 
                    set_engine("glm")

# Modelling workflow for pairing model and recipe 
flu_mod10_workflow2 <- workflow() %>% 
  add_model(flu_mod10_mod2) %>% 
  add_recipe(flu_mod10_rec2)

flu_mod10_workflow2
# Using the resulting predictors for preparing recipe and training the model
flu_mod10_fit2 <- 
 flu_mod10_workflow2 %>% 
  fit(data = train_data)

# Pulling the fitted model object and using tidy() function for getting a tidy tibble of model coefficients
flu_mod10_fit2 %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

**ALTERNATIVE MODEL: USING TRAINED WORKFLOW TO PREDICT**

```{r, warning = FALSE, message=FALSE}
# Using the trained data to predict with the unseen test data
predict(flu_mod10_fit2, test_data)

# Using argument() with the model plus test data for saving them together
flu_mod10_aug2 <- 
  augment(flu_mod10_fit2, test_data)
  
flu_mod10_aug2 %>%
  select(Nausea, .pred_No, .pred_Yes)
  
# Creating ROC curve and piping to the autoplot() method
flu_mod10_aug2 %>% 
  roc_curve(truth = Nausea, .pred_No) %>% 
  autoplot()
# Estimating the area under the curve 
flu_mod10_aug2 %>% 
  roc_auc(truth = Nausea, .pred_No)
```

Given this result, the full model using all variables as predictors perform better the reduced model with only RunnyNose as predictor. The full model as a ROC-AUC of 0.71 compared the reduced model with ROC-AUC of 0.46

