---
title: "Module 8: Data Wrangling - Chris Okitondo"
output:
  html_document:
    toc: FALSE
---

### Setup
Load needed packages. 
```{r, echo=FALSE, message=FALSE}
library('here')
library('tidyverse')
library('dplyr')
```


### Data loading
```{r}
# path to data
data_location<- here::here("fluanalysis", "data", "SympAct_Any_Pos.Rda")
# load data
data <- readRDS(data_location)
```


### Looking at the data
```{r}
dplyr::glimpse(data)
summary(data)
head(data)
skimr::skim(data)
```

### Checking variable names 
```{r}
names(data)
```

### Step 1: Removing all variables that have Score or Total or FluA or FluB or Dxname or Activity in their name
```{r}
data1 <- data %>% dplyr:: select(-matches("Score|Total|FluA|FluB|Dxname|Activity"))
```


### Step 2: Removing the variable Unique.Visit.
```{r}
data2 <- data1[,-1]
```

### Step 3: Removing all NA observations in the dataset
```{r}
data3 <- na.omit(data2)
```

### Naming the newly processed data as cleaned_data
```{r}
cleaned_data <- data3
```

### Saving the cleaned data
```{r}
save_data_location <- here::here("fluanalysis", "data", "cleaned_data.rds")
saveRDS(cleaned_data, file = save_data_location)
```


### Addition for Module 11: Preparing data for Machine Learning


#### Loading the previous cleaned data for additional cleaning
```{r}
cleaned_data <- readRDS(here("fluanalysis", "data", "cleaned_data.rds"))
```


#### We are getting rid of repetitive variables
```{r}
cleaned_data_v2 <- cleaned_data %>% select(-CoughYN,-WeaknessYN,-CoughYN2,-MyalgiaYN)

```

#### We are getting rid of hearing and vision variables since they have less than 5- entries (Y/N)
```{r}
cleaned_data_v3 <- cleaned_data_v2 %>% select(-Vision,-Hearing)
```

Our data is now made of 730 observations and 26 variables. This is what we will use for Module 11.
```{r}
saveRDS(cleaned_data_v3, file = here("fluanalysis", "data", "cleaned_data_mod11.rds"))
```






